Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2017-09-29 paper/aldebaran.tex
tidy introduction

Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations. It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion \citep[the radial velocity or RV method;][]{struverv}. While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran. After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} now claim a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new RV observations with the SONG~Telescope, and photometry from the \ktwo Mission. By measuring the frequency of maximum power ,\numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun. This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.

>>>>

Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations. It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion \citep[the radial velocity or RV method;][]{struverv}. While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran. After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission. By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun. This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
